Visualizing probability distributions across bivariate cyclic temporal granularities

### 2021-04-22 main.Rmd
comments update
definition without section

bibliography: [bibliography.bib]
preamble: >
  \usepackage{mathtools,amssymb,booktabs,amsthm,todonotes,colortbl}
  \def\mod{~\text{mod}~}
  \newtheorem{definition}{Definition}
  \usepackage{mathptmx}
  \usepackage{caption}
  \DeclareCaptionStyle{italic}{labelfont={bf},textfont={it},labelsep=colon}
  \captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
  \captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
  \def\novspacing{\setlength{\aboverulesep}{0pt}\setlength{\belowrulesep}{0pt}}
  \def\vspacing{\setlength{\aboverulesep}{0.4ex}\setlength{\belowrulesep}{0.65ex}}
output:
  bookdown::pdf_book:
    base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
---

@@

\begin{definition}\label{def:definition}
A \textbf{time domain} is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
\end{definition}

@@

\begin{definition}\label{def:index set}
The \textbf{index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, uniquely maps the time instants to the set of non-negative integers.
\end{definition}

\begin{definition}\label{def:linear}
A \textbf{linear granularity} is a mapping $G$ from the index set, $Z$, to subsets of the time domain such that:
  (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less than all elements of $G(j)$; and
  (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
Each non-empty subset $G(i)$ is called a \textbf{granule}.
\end{definition}

@@

\begin{definition}\label{def:finerthan}
A linear granularity $G$ is \textbf{finer than} a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that
$G(i) \subset H(j).$
\end{definition}

\begin{definition}\label{def:groupsinto}
A linear granularity $G$ \textbf{groups into} a linear granularity $H$, denoted
$G \trianglelefteq H$, if for each index $j$ there exists a (possibly infinite) subset $S$ of the integers such that $H(j) = \bigcup_{i \in S}G(i).$
\end{definition}

@@

\begin{definition}\label{def:periodic}
A granularity $G$ is \textbf{periodic} with respect to a finite granularity $H$ if:
(1) $G \trianglelefteq H$; and
(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z_\geq 0}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
\end{definition}

@@

\begin{definition}\label{def:order}
The \textbf{order} of a linear granularity is the level of coarseness associated with a linear granularity. A linear granularity G will have lower order than H if each granule of G is composed of lower number of granules of bottom granularity than each granule of H.
\end{definition}

@@

\begin{definition}\label{def:circular}
A \textbf{circular granularity} $C_{B, G}$ relates linear granularity $G$ to bottom granularity $B$ if
\begin{equation} \label{eq:circular-gran}
\begin{split}
C_{B, G}(z) & = z\mod P(B, G) \quad \forall z \in \mathbb{Z}_{\geq 0} \\
\end{split}
\end{equation}
where
$z$ denotes the index set,
$B$ groups periodically into $G$ with regular mapping and period $P(B, G)$.
\end{definition}

@@

\begin{definition}\label{def:quasicircular}
A \textbf{quasi-circular granularity} $Q_{B, G'}$ is formed when bottom granularity $B$ groups periodically into linear granularity $G'$ with irregular mapping such that the granularities are given by
\begin{equation}\label{eq:quasi}
Q_{B, G'}(z) =
z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert, \quad \text{for}\quad z \in T_{k},
\end{equation}
where
$z$ denotes the index set,
$w$ denotes the index of $G'$,
$R'$ is the number of granules of $G'$ in each period of $(B, G')$,
$T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$,
and $\vert T_w \vert$ is the cardinality of set $T_w$.
\end{definition}

@@

\begin{definition}\label{def:aperiodic}
An \textbf{aperiodic cyclic granularity} is formed when bottom granularity $B$ groups into an aperiodic linear granularity $M$ such that the granularities are given by
\begin{equation}\label{eq:aperiodic}
A_{B, M}(z) = \begin{cases}
                  i, & \text{for}\quad z \in T_{i_j} \\
                  0  & \text{otherwise},
                \end{cases}
\end{equation}
where
$z$ denotes the index set,
$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$, $n$ being the number of  aperiodic linear granularities in consideration.
\end{definition}

@@

\begin{definition}\label{def:hierarchy}
The nested ordering of linear granularities can be organized into a \textbf{hierarchy table}, denoted as $H_n: (G, C, K)$, which arranges them from lowest to highest in order. It shows how the $n$ granularities relate through $K$, and how the cyclic granularities, $C$, can be defined relative to the linear granularities. Let $G_{\ell}$ and $G_{m}$ represent the linear granularity of order $\ell$ and $m$ respectively with $\ell<m$. Then $K \equiv P(\ell,m)$ represents the period length of the grouping $(G_{\ell}, G_{m})$, if $C_{G_{\ell}, G_{m}}$ is a circular granularity and $K \equiv k(\ell,m)$ represents the operation to obtain $G_{m}$ from $G_{\ell}$, if $C_{G_{\ell}, G_{m}}$ is quasi-circular.
\end{definition}

@@

\begin{definition}\label{def:norderup}
The hierarchy table contains \textbf{multiple-order-up} granularities which are cyclic granularities that are nested within multiple levels.
A \textbf{single-order-up} is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
\end{definition}

@@

\begin{definition}\label{def:clash}
A \textbf{clash} is a pair of cyclic granularities that contains empty combinations of categories.
\end{definition}

\begin{definition}\label{def:harmony}
A \textbf{harmony} is a pair of cyclic granularities that does not contain any empty combinations of its categories.
\end{definition}

>>>>

bibliography: [bibliography.bib]
preamble: >
  \usepackage{mathtools,amssymb,booktabs,amsthm,todonotes,colortbl}
  \def\mod{~\text{mod}~}
  \usepackage{mathptmx}
  \usepackage{caption}
  \newtheorem{defn}{Definition}
  \DeclareCaptionStyle{italic}{labelfont={bf},textfont={it},labelsep=colon}
  \captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
  \captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
  \def\novspacing{\setlength{\aboverulesep}{0pt}\setlength{\belowrulesep}{0pt}}
  \def\vspacing{\setlength{\aboverulesep}{0.4ex}\setlength{\belowrulesep}{0.65ex}}
output:
  bookdown::pdf_book:
    base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
---

@@

\begin{defn}\label{def:timedom}
A \textbf{time domain} is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
\end{defn}

@@

\begin{defn}\label{def:index set}
The \textbf{index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, uniquely maps the time instants to the set of non-negative integers.
\end{defn}

\begin{defn}\label{def:linear}
A \textbf{linear granularity} is a mapping $G$ from the index set, $Z$, to subsets of the time domain such that:
  (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less than all elements of $G(j)$; and
  (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
Each non-empty subset $G(i)$ is called a \textbf{granule}.
\end{defn}

@@

\begin{defn}\label{def:finerthan}
A linear granularity $G$ is \textbf{finer than} a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that
$G(i) \subset H(j).$
\end{defn}

\begin{defn}\label{def:groupsinto}
A linear granularity $G$ \textbf{groups into} a linear granularity $H$, denoted
$G \trianglelefteq H$, if for each index $j$ there exists a (possibly infinite) subset $S$ of the integers such that $H(j) = \bigcup_{i \in S}G(i).$
\end{defn}

@@

\begin{defn}\label{def:periodic}
A granularity $G$ is \textbf{periodic} with respect to a finite granularity $H$ if:
(1) $G \trianglelefteq H$; and
(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z_\geq 0}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
\end{defn}

@@

\begin{defn}\label{def:order}
The \textbf{order} of a linear granularity is the level of coarseness associated with a linear granularity. A linear granularity G will have lower order than H if each granule of G is composed of lower number of granules of bottom granularity than each granule of H.
\end{defn}

@@

\begin{defn}\label{def:circular}
A \textbf{circular granularity} $C_{B, G}$ relates linear granularity $G$ to bottom granularity $B$ if
\begin{equation} \label{eq:circular-gran}
\begin{split}
C_{B, G}(z) & = z\mod P(B, G) \quad \forall z \in \mathbb{Z}_{\geq 0} \\
\end{split}
\end{equation}
where
$z$ denotes the index set,
$B$ groups periodically into $G$ with regular mapping and period $P(B, G)$.
\end{defn}

@@

\begin{defn}\label{def:quasicircular}
A \textbf{quasi-circular granularity} $Q_{B, G'}$ is formed when bottom granularity $B$ groups periodically into linear granularity $G'$ with irregular mapping such that the granularities are given by
\begin{equation}\label{eq:quasi}
Q_{B, G'}(z) =
z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert, \quad \text{for}\quad z \in T_{k},
\end{equation}
where
$z$ denotes the index set,
$w$ denotes the index of $G'$,
$R'$ is the number of granules of $G'$ in each period of $(B, G')$,
$T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$,
and $\vert T_w \vert$ is the cardinality of set $T_w$.
\end{defn}

@@

\begin{defn}\label{def:aperiodic}
An \textbf{aperiodic cyclic granularity} is formed when bottom granularity $B$ groups into an aperiodic linear granularity $M$ such that the granularities are given by
\begin{equation}\label{eq:aperiodic}
A_{B, M}(z) = \begin{cases}
                  i, & \text{for}\quad z \in T_{i_j} \\
                  0  & \text{otherwise},
                \end{cases}
\end{equation}
where
$z$ denotes the index set,
$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$, $n$ being the number of  aperiodic linear granularities in consideration.
\end{defn}

@@

\begin{defn}\label{def:hierarchy}
The nested ordering of linear granularities can be organized into a \textbf{hierarchy table}, denoted as $H_n: (G, C, K)$, which arranges them from lowest to highest in order. It shows how the $n$ granularities relate through $K$, and how the cyclic granularities, $C$, can be defined relative to the linear granularities. Let $G_{\ell}$ and $G_{m}$ represent the linear granularity of order $\ell$ and $m$ respectively with $\ell<m$. Then $K \equiv P(\ell,m)$ represents the period length of the grouping $(G_{\ell}, G_{m})$, if $C_{G_{\ell}, G_{m}}$ is a circular granularity and $K \equiv k(\ell,m)$ represents the operation to obtain $G_{m}$ from $G_{\ell}$, if $C_{G_{\ell}, G_{m}}$ is quasi-circular.
\end{defn}

@@

\begin{defn}\label{def:norderup}
The hierarchy table contains \textbf{multiple-order-up} granularities which are cyclic granularities that are nested within multiple levels.
A \textbf{single-order-up} is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
\end{defn}

@@

\begin{defn}\label{def:clash}
A \textbf{clash} is a pair of cyclic granularities that contains empty combinations of categories.
\end{defn}

\begin{defn}\label{def:harmony}
A \textbf{harmony} is a pair of cyclic granularities that does not contain any empty combinations of its categories.
\end{defn}
